### 2025-06-06 paper/thesis.tex

\documentclass[a4paper,12pt]{report}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins according to Bonn University guidelines
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=2cm]{geometry}

%% Font settings - Times New Roman or similar (as per guidelines)
\usepackage{times} % Times font
\usepackage{mathptmx} % Times font with math support

%% Section formatting
\usepackage{titlesec}

% Traditional academic chapter title style
\titleformat{\chapter}[display]
  {\normalfont\Large\bfseries}
  {}
  {0pt}
  {\thispagestyle{empty}%
   \vspace*{-20pt}%
   \begin{flushright}
     % Elegant, academic chapter number
     \begin{tikzpicture}[remember picture, overlay]
       \node[text=black!12, font=\fontsize{80}{80}\selectfont] {\thechapter};
     \end{tikzpicture}
   \end{flushright}
   \vspace{-60pt}
   % Simple, elegant single line
   \rule{\textwidth}{0.6pt}\\
   \vspace*{15pt}
   \raggedright\LARGE\scshape}
  [\vspace{-0.8cm}]
\titlespacing*{\chapter}{0pt}{0pt}{40pt}

% Elegant section formatting
\titleformat{\section}
  {\normalfont\Large\bfseries\scshape}
  {}
  {0pt}
  {\thesection\quad}
\titlespacing*{\section}{0pt}{3.5ex plus 1ex minus .2ex}{1.5ex plus .2ex}

% Make subsections italic and elegant
\titleformat{\subsection}
  {\normalfont\large\bfseries\itshape}
  {\thesubsection}
  {1em}
  {}
\titlespacing*{\subsection}{0pt}{3ex plus 1ex minus .2ex}{1.5ex plus .2ex}

%% Line spacing 1.5 as required
\usepackage{setspace}
\onehalfspacing

%% Useful packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}         % For bold symbols
\usepackage{graphicx}
\usepackage{float}      % For [H] figure placement
\usepackage{booktabs}
\usepackage{tabu}
\usepackage{enumitem}
\usepackage[table,dvipsnames]{xcolor}  % Added dvipsnames option for more color choices
\usepackage{array}
\usepackage{multirow}
\usepackage{amsthm}
\usepackage{bbm}        % Blackboard math symbols (for indicator function)
\usepackage[toc]{appendix}
\usepackage{subcaption}

% Choose either algorithm2e OR algorithmicx, not both
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

% For adjusting margins of specific sections
\usepackage{changepage}

% Bibliography
\usepackage[natbibapa]{apacite}

% Todo notes
\usepackage[colorinlistoftodos]{todonotes}

% Page numbering configuration
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}

% Hyperref should be last
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

% Allow floating content to be placed more liberally
\renewcommand{\floatpagefraction}{.9}
\renewcommand{\textfraction}{.1}
\renewcommand{\topfraction}{0.9}
\renewcommand{\bottomfraction}{.9}

% Appendix formatting
\makeatletter
\let\old@chapter\@chapter
\def\@chapter[#1]#2{\old@chapter[{#1}]{#2}}
\makeatother

% Math operators
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% Custom theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

% Define indicator function notation
\newcommand{\ind}[1]{\mathbbm{1}_{#1}}

\begin{document}

% Cover sheet as per Bonn University guidelines
\begin{titlepage}
\begin{center}
\vspace*{2cm}

{\Large \textbf{Social Learning via\\[0.25cm]Multi-Agent Deep Reinforcement Learning}}\\[6cm]

Master Thesis Presented to the\\
Department of Economics at the\\
Rheinische Friedrich-Wilhelms-Universität Bonn\\[1cm]

In Partial Fulfillment of the Requirements for the Degree of\\
Master of Science (M.Sc.)\\[4cm]

Supervisor: Prof. Dr. Florian Brandl\\[2cm]

Submitted in June, 2025 by:\\
Ege Can Doğaroğlu\\
Matriculation Number: 3464688

\end{center}
\end{titlepage}

% Set page numbering to 'gobble' (no numbers) for front matter
\pagenumbering{gobble}

% Abstract
\begin{abstract}
\begin{adjustwidth}{2cm}{2cm}
This thesis bridges economic social learning theory and multi-agent reinforcement learning by developing Partially Observable Active Markov Games (POAMGs) and the POLARIS algorithm. Our framework addresses fundamental challenges in modeling adaptive behavior under partial observability, where agents must simultaneously learn about uncertain environments and evolving strategies of others. We extend Active Markov Games to partially observable settings, derive theoretical convergence results, and develop a practical algorithm combining Transformer-based belief processing, Graph Neural Networks with attention mechanisms, and reinforcement learning optimization. Applied to strategic experimentation and learning without experimentation scenarios, our approach reveals dynamic role assignment as a robust organizing principle where agents naturally differentiate into complementary roles enhancing collective information processing. We find that free-riding behavior does not lead to uniform inefficiencies—for some agents, performance in larger networks exceeds autarky levels. Our two-staged analysis methodology addresses computational challenges like catastrophic forgetting while providing systematic approaches for extracting insights from multi-agent learning dynamics. The framework contributes theoretical foundations to multi-agent reinforcement learning and computational tools to economic theory, demonstrating how interdisciplinary approaches can illuminate complex social learning phenomena neither field could address alone.
\end{adjustwidth}
\end{abstract}

% Table of contents
\tableofcontents
\newpage

% Optional - include if you have abbreviations, tables, or figures to list
%\listoffigures
%\listoftables
%\newpage

% Page numbering starts here with Arabic numerals
\chapter{Introduction}
\pagenumbering{arabic}
\setcounter{page}{1}
\input{intro}
\chapter{Related Work}
\input{background}
\chapter{Reinforcement Learning Framework}
\input{theory}
\chapter{Evaluation}
\input{results}
\chapter{Conclusion}
\input{conclusion}

% Appendix
\newpage
\appendix
\addcontentsline{toc}{chapter}{Appendices}
\input{appendix}

% Bibliography
\newpage
\renewcommand{\bibname}{References}
\renewcommand{\thechapter}{R}
\addcontentsline{toc}{chapter}{References}
\bibliographystyle{apacite}
\bibliography{refs}
\renewcommand{\thechapter}{\arabic{chapter}} % Reset chapter numbering

% Statement of Authorship (as required)
\newpage
\section*{Statement of Authorship}

I hereby confirm that the work presented has been performed and interpreted solely by myself except for where I explicitly identified the contrary. I assure that this work has not been presented in any other form for the fulfillment of any other degree or qualification. Ideas taken from other works in letter and in spirit are identified in every single case.\\[2cm]

\noindent Date: \rule{4cm}{0.5pt} \hfill Signature: \rule{6cm}{0.5pt}

\end{document}